\documentclass[letterpaper,11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{times}
\usepackage[compact]{titlesec}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage[T1]{fontenc}
\usepackage{biblatex}
\usepackage{tabularx}
\usepackage{cleveref}
\usepackage{tikz}
\usetikzlibrary{arrows.meta}
\usetikzlibrary{shapes.geometric}
\usetikzlibrary{shapes.multipart}
\usetikzlibrary{cd}

\addbibresource{ref23.bib}

\graphicspath{
    {Figs/}
}

\begin{document}

\fontfamily{cmss}\selectfont
\thispagestyle{empty}
\noindent\textbf{\Large How To Contribute to the Annual Report}\\[0.3cm]

This document describes how each co-PI should contribute to the annual report. The official AFOSR guideline is available \href{https://afosr.gov1.qualtrics.com/CP/File.php?F=F_6z0CePSAlQtlzng}{here}. 

\begin{enumerate}
    \item Enter project data
        \begin{itemize}
            \item Go to the google sheet, \href{https://docs.google.com/spreadsheets/d/10obUfCgK5TLNYDSzH0GSxBX0Bv0RVghhWma4WrY66xU/edit?usp=sharing}{\texttt{MURI24 report data}}, where the data from the last year has been copied to be edited by you
            \item At the bottom, click the tab of \texttt{Participants}, and list yourself and anyone participated in your work. 
            \item At the bottom, click the next tab of \texttt{Collaboration}, and enter the info for your partner/collaborating organizations. 
            \item At the bottom, click the next tab of \texttt{Intellectual Property}, and mark any product, such as patent or license. 
        \end{itemize}
    \item Enter publications
        \begin{itemize}
            \item Go to the overleaf project, \href{https://www.overleaf.com/1586261996hchzjbrvfgmj#6d8639}{\texttt{muri\_report23}}
            \item Add your publications to the bib file,  \texttt{ref23.bib}
            \item If available, please add the DOI number under the \texttt{doi} field, e.g., \texttt{doi = \{10.1.1/jpb001\}}
            \item It is being assumed that every publication includes the acknowledgment of the AFOSR support. Let Taeyoung know otherwise.
        \end{itemize}
    \item Complete a section or more for accomplishments
        \begin{itemize}
            \item Go to the overleaf project, \href{https://www.overleaf.com/1586261996hchzjbrvfgmj#6d8639}{\texttt{muri\_report23}}
            \item At the latex file \texttt{report23.tex}, find the subsection \texttt{Topic ?}. Change the title of the subsection into the topic of your investigation with the last name(s) of the participating PI(s) in parenthesis, e.g., \textbf{Stochastic hybrid system (Lee, Clark)}
            \item You may choose one or more topics that you worked on. 
            \item For each subsection corresponding to a single topic, add the description of your accomplishments according to the following guide:
                \begin{itemize}
                    \item Research objectives: list the major goals of your work and describe the proposed technical approach to achiee those goals.
                    \item Accomplishments: describe major activities, significant results or key outcomes, including major findings, developments, or positive/negative conclusions. 
                    \item Plan: summarize the objectives to be achieved for the next reporting period. 
                \end{itemize}
            \item Since we are going to cite our papers, these descriptions do not have to include all the technical details. 
            \item It is suggested that each subsection is about 0.75-1 page. 
        \end{itemize}
\end{enumerate}

Please, submit your input by \textbf{Friday, July 19th}. 
As these are completed by each PI, Taeyoung will work on the remaining parts of the report. He also may reorganize your section to make the report more streamlined and coherent.  

\end{document}
